|

### 2020-04-16 paper.Rmd
Keeping tex file for submission.
Preparing for resubmission.

---
title: |
  A slice tour for finding hollowness in high-dimensional data
type: Short Technical Note
author:
  - name: Ursula Laa
    affil: a, b
    email: ursula.laa@monash.edu
  - name: Dianne Cook
    affil: b
    email: dicook@monash.edu
  - name: German Valencia
    affil: a
    email: german.valencia@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: |
  Taking projections of high-dimensional data is a common analytical and visualisation technique in statistics for working with high-dimensional problems. Sectioning, or slicing, through high dimensions is less common, but can be useful for visualising data with concavities, or non-linear structure. It is associated with conditional distributions in statistics, and also linked brushing between plots in interactive data visualisation. This short technical note describes a simple approach for  slicing in the orthogonal
  space of projections obtained when running a tour, thus presenting the viewer
  with an interpolated sequence of sliced projections. The method has been
  implemented in R as an extension to the tourr package, and can be used to
  explore for concave and non-linear structures in multivariate distributions.
keywords: |
  data visualisation; grand tour; sectioning; statistical computing; statistical graphics; high-dimensional data
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
---

@@

<!-- The disadvantage of projection: opacity, occlusion, even with pseudo-transparency from scatterplots;  motivation -->
A major limitation of projections is their opacity. It is mitigated some, when scatterplots are used to render the projected data, through a pseudo-transparency of sparseness of points. This is improved further if points are also drawn using an alpha level that provides transparent dots on most display devices. Some features of multivariate distributions may be visible, but a lot is easily hidden, especially in the case of concave structures. Considering the example of a simple geometric shape such as a hypershere, it is difficult to distinguish between a full or a hollow sphere, that is, whether points are uniformly distributed within the sphere or on the surface of a sphere. Similarly we can think of small scale structures hidden in the centre of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections. Projections also obscure non-linear boundaries as might be constructed from a classification model, or nonlinear model fits in high-dimensions.

<!-- Our approach: introduce sections; many different ways to slice, refer to linked brushing as slicing whcih mirrors prosection, but here we will explain/utilise a simple slicing on the projection where point in orthogonal space are faded. -->
Slicing, or sectioning, is a way that the internal distribution of high-dimensional data can be explored. @prosection discusses a technique for combining projections with sections constructed by slicing in the dimensions orthogonal to the projection. It is also possible to think of linked brushing (e.g. @ggobiCSDA2003, @JSSv081i05) as slicing. In this note, we discuss an approach to sectioning, where observations in the space orthogonal to the projection are highlighted if they are close to a projection plane through the mean of the data, and faded if further afield. This is combined with the grand tour to provide a new dynamic display that can be used to systematically search for features hidden in high dimensions.

@@

A tour provides a continuous sequence of $d$-dimensional (typically $d=1$ or $2$) projections from $p$-dimensional Euclidean space. It is constructed by combining a method for basis selection with geodesic interpolation between pairs of bases. In a grand tour, the basis selection is random, each new basis is chosen from all possible projections. In a guided tour, the bases are chosen based on an index of interestingness. Different basis selection methods as well as the geodesic interpolation are implemented in the `tourr` package [@tourr], which also provides several display functions for viewing the tour.

For the explanation of the slice tour, the actual mechanics of the tour are not important, and only the notion of a projection plane in high dimensions is needed. The notation used for explanations starts with denoting each projection as $\mathbf{X}$ the $n\times p$ dimensional data matrix ($n$ observations in $p$ dimensions) and $\mathbf{A}$ an orthonormal $p \times d$ projection matrix. The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, producing the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.

@@

The orthogonal distance of the data points from the current projection plane is calculated as,

\begin{equation}
\tilde{v}_i^2 = ||\mathbf{x}_i'||^2,
\end{equation}

\noindent where

\begin{equation}
\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2
\end{equation}

\noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$. $\mathbf{A}$ is assumed to be through 0, because the data has been centred on its mean. $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane. This can be generalized for $d>2$.

The distance can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$. In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$. We could use this single direction to slice systematically from one side of the space to the other. It is also possible to think of front and back, some points are closer to the viewer, and some are far, and depth cues could be used. When going beyond 3D, the dimension of the orthogonal space is $>1$, and all sense of direction is lost. In general, there are $(p-d)$ dimensions orthogonal to the projection plane, and using distance from the points to the plane is the simplest approach to generating slices. Using the Euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.

@@

The fraction of points inside a slice of thickness $h$ can be estimated as the relative volume of the slice compared to the volume of the full hypersphere,
\begin{equation}
V_{rel} = \frac{1}{2} \frac{h^{p-2}}{R^p} (p R^2 - (p-2) h^2) \approx \frac{1}{2} (\frac{h}{R})^{p-2},
\end{equation}
where $R$ is the radius of the hypersphere. The approximation is valid when $h \ll R$, as is typically expected to be the case. To keep the relative number of points (i.e. $V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $\epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.1.

### Non-central slice

The equations above assume that the projection plane, and thus the the slice, passes through 0, which will be the mean of centred data. This (centre point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too. However it is straightforward to generalize the equations to use any centre point, $\mathbf{c}$.

@@

The slice tour has been implemented in R [@rref] as a new display method `display_slice` in the `tourr` package [@tourr]. In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the centre point $\mathbf{c}$ is set by the `anchor` argument. By default `eps=0.1` and `anchor=NULL`, resulting in slicing through the mean. In addition the user can select the marker symbols for point inside and outside the slice. By default points in the slice are highlighted as `pch_slice = 20` and `pch_other = 46`, i.e. plotting a bullet for points inside the slice, and a dot for points outside. Below we show example code for displaying slices through a hollow 3D sphere.

@@

It is especially instructive to look at spheres that are sliced off-centre. In this case the different views obtained in the slice tour reveal more of the concave structure of the distribution. The first row of Figure \ref{fig:3anchored} shows example views from a slice tour on a 3D sphere with $R=1$ and centre point $(0.7, 0.7, 0.7)$. Notice that this was chosen to fall outside the sphere. Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $R$ (when the centre point has a negligible orthogonal component to the viewing plane); a circle with radius $<R$ as the viewing angle is tilted away from the axis connecting the centre point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the centre point axis. A similar picture is found for the 5D sphere, see second row of Figure \ref{fig:3anchored}. For this example we generate 20k points on a hollow 5D sphere to resolve the features. Since $h$ increases with $p$ the resolution is reduced compared to the 3D example. Note also that as dimensionality increases, the larger orthogonal space means that the centre point will have a large orthogonal component in most views.

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored/sphere-3-anchored-033.png}
\hspace{5mm}
\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored/sphere-3-anchored-036.png}
\hspace{5mm}
\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored/sphere-3-anchored-041.png}}
\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-014.png}
\hspace{5mm}
\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-010.png}
\hspace{5mm}
\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-001.png}}
\caption{Different slices through a 3D (first row) and 5D (second row) hollow sphere with $R=1$, with shifted centre point, showing full circle with small radius (left), circle with radius $<R$ (middle) and circle with radius $R$ (right).}
\label{fig:3anchored}
\end{figure*}

### Other geometric shapes
To better understand the slice tour we look at different examples of geometric shapes, see Figure \ref{fig:geoms}. For each shape two selected views are shown. The first column shows views from the slice tour on a 3D Roman surface. The second column shows slices through a 4D torus, revealing different aspects of the shape. The last column shows a 6d cube, where the upper plot shows a view along two of the original parameters, allowing to clearly identify the rectangular shape. The panel below shows that this is not typically the case when looking at a randomly selected slice.

@@

Selected views of projections and sliced projections from a radial basis SVM on 3 variables, and a polynomial basis SVM on 5 variables are shown in Figure \ref{fig:wine}. The slicing allows exploring the centre of the space. With 3D it reveals the spherical boundary of the group (green) hidden by the projection. In 5D with the polynomial basis, projection might suggest that the boundary between classes is almost linear, but the slicing shows that it to be nonlinear near the centre.

\begin{figure*}[ht]
\centerline{
\includegraphics[width=0.3\textwidth]{gifs/png/wine-radial-projected/wine-radial-projected-031.png}
\hspace{5mm}
\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly-projected/wine-poly-projected-096.png}
\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly-projected/wine-poly-projected-042.png}}
\centerline{
\includegraphics[width=0.3\textwidth]{gifs/png/wine-radial/wine-radial-031.png}
\hspace{5mm}
\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly/wine-poly-096.png}
\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly/wine-poly-042.png}
}
\caption{Exploring classification boundaries of the wine dataset using projection (top row) and sliced projection (bottom row). Radial basis SVM (first column) of 3 variables shows how the slicing reveals the spherical shape of the boundary of one group (green) that was hidden in the projection. Polynomial basis SVM (second and third columns) of 5 variables. The orange group does tend to be wrapped by the green group, and the blue group disappears with slicing, showing that it is on the outer edge of the space.}
\label{fig:wine}
\end{figure*}

@@

This paper has introduced a new visualization method for dynamic slicing of high-dimensional spaces. It is based on interpolated projections obtained in a (grand) tour and generates an interpolated sequence of sliced projections. The examples shown in Section \ref{sec:examples} demonstrate the potential of this new display to find and explore concave structures, as well as other hidden features.

@@

The approach by @prosection is more complex, and slices a subspace  of the $p-d=p-2$ dimensional space orthogonal to the projection. This generates a more parameters, making it more difficult to navigate. More parameters mean more decisions on what to show. However, it is one of the next steps to explore different definitions of a slice tour, and the types of structure that might be captured by variations in slicing algorithms.

@@

The authors gratefully acknowledge the support of the Australian Research Council. The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr]. The source material and animated gifs for this paper are available at \url{https://github.com/uschiLaa/paper-slice-tour}. An appendix deriving the relative slice volume for the hypersphere is included as supplemental material.

>>>>

---
title: |
  A slice tour for finding hollowness in high-dimensional data
type: Short Technical Note
author:
  - name: Ursula Laa
    affil: a, b
    email: ursula.laa@monash.edu
  - name: Dianne Cook
    affil: b
    email: dicook@monash.edu
  - name: German Valencia
    affil: a
    email: german.valencia@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: |
  Taking projections of high-dimensional data is a common analytical and visualisation technique in statistics for working with high-dimensional problems. Sectioning, or slicing, through high dimensions is less common, but can be useful for visualising data with concavities, or non-linear structure. It is associated with conditional distributions in statistics, and also linked brushing between plots in interactive data visualisation. This short technical note describes a simple approach for  slicing in the orthogonal
  space of projections obtained when running a tour, thus presenting the viewer
  with an interpolated sequence of sliced projections. The method has been
  implemented in R as an extension to the tourr package, and can be used to
  explore for concave and non-linear structures in multivariate distributions.
keywords: |
  data visualisation; grand tour; sectioning; statistical computing; statistical graphics; high-dimensional data
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
keep_tex: true
---

@@

<!-- The disadvantage of projection: opacity, occlusion, even with pseudo-transparency from scatterplots;  motivation -->
A major limitation of projections is their opacity. It is mitigated somewhat when scatterplots are used to render the projected data, through a pseudo-transparency of sparseness of points. This is improved further if points are also drawn using an alpha level that provides transparent dots on most display devices. Some features of multivariate distributions may be visible, but a lot is easily hidden, especially in the case of concave structures. Consider the example of a simple geometric shape such as a hypersphere. It is difficult to distinguish between a full or a hollow sphere, that is, whether points are uniformly distributed within the sphere or on the surface of a sphere.\footnote{\textcolor{red}{In standard terms of geometry we are comparing a disk (or ball), $D^p = \{ x \in \mathbb{R}^p : |x| \leq 1\}$ with a standard sphere, $S^{p-1} = \{ x \in \mathbb{R}^p : |x| = 1\}$.}} Similarly we can think of small scale structures hidden in the centre of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections. Projections also obscure non-linear boundaries as might be constructed from a classification model, or nonlinear model fits in high-dimensions.

<!-- Our approach: introduce sections; many different ways to slice, refer to linked brushing as slicing whcih mirrors prosection, but here we will explain/utilise a simple slicing on the projection where point in orthogonal space are faded. -->
Slicing, or sectioning, is a way that the internal distribution of high-dimensional data can be explored. \textcolor{red}{Here we use "slice" and "section" interchangably, to refer to a portion of the full parameter space spanned by the data distribution. The section is defined by limiting the parameter ranges, thus selecting only a sub-space, and in general will be determined by a set of inequality conditions. Using a linear condition for going from 3D to 2D we can think of flat slices through the space.} @prosection discusses a technique for combining projections with sections constructed by slicing in the dimensions orthogonal to the projection. It is also possible to think of linked brushing (e.g. @ggobiCSDA2003, @JSSv081i05) as slicing. In this note, we discuss an approach to sectioning, where observations in the space orthogonal to the projection are highlighted if they are close to a projection plane through the mean of the data, and faded if further afield. This is combined with the grand tour to provide a new dynamic display that can be used to systematically search for features hidden in high dimensions.

@@

A tour provides a continuous sequence of $d$-dimensional (typically $d=1$ or $2$) projections from $p$-dimensional Euclidean space. It is constructed by combining a method for basis selection with geodesic interpolation between pairs of bases. In a grand tour, the basis selection is random; each new basis is chosen from all possible projections. In a guided tour, the bases are chosen based on an index of interestingness. Different basis selection methods as well as the geodesic interpolation are implemented in the `tourr` package [@tourr], which also provides several display functions for viewing the tour.

For the explanation of the slice tour, the actual mechanics of the tour are not important, and only the notion of a projection plane in high dimensions is needed. The data consists of $n$ observations in $p$-dimensions ($n$ points in $\mathbb{R}^p$) and is thus represented as an $n\times p$- dimensional data matrix $\mathbf{X}$. We denote by $\mathbf{A}$, a $p \times d$-dimensional (projection) matrix, the columns of which span a $d$-dimensional subspace of $\mathbb{R}^p$. The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, producing the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.

@@

The orthogonal distance of the data points from the current projection plane is calculated as, 
\begin{equation}
\tilde{v}_i^2 = ||\mathbf{x}_i'||^2,
\end{equation}
\noindent where 
\begin{equation}
\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2
\end{equation}
\noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$. \textcolor{red}{The plane defined by $\mathbf{A}$ passes through 0, suggesting that the data should be centred on its mean. Alternatively, we can use a rigid translation to a parallel affine plane that passes through the data mean, as discussed in Section \ref{sec:offcentre}.} $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane. This can be generalized for $d>2$.

The distance can then be used to display a slice tour by highlighting \textcolor{red}{the projected} points for which the orthogonal distance is smaller than a selected cutoff value $h$. In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$. We could use this single direction to slice systematically from one side of the space to the other, as illustrated in Figure \ref{fig:diagrams}. It is also possible to think of front and back; some points are closer to the viewer, and some are far, and depth cues could be used. \textcolor{red}{When the orthogonal dimension $p-d \geq 2$, there is no longer a unique orthogonal direction in which to translate the projection plane. In general,} using distance from the points to the plane is the simplest approach to generating slices. Using the Euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.

@@

The fraction of points inside a slice of thickness $h$ can be estimated as the relative volume of the slice compared to the volume of the full hypersphere,
\begin{equation}
V_{rel} = \frac{1}{2} \frac{h^{p-2}}{R^p} \left(p R^2 - (p-2) h^2\right) \approx \frac{1}{2} \left(\frac{h}{R}\right)^{p-2},
\end{equation}
where $R$ is the radius of the hypersphere. The approximation is valid when $h \ll R$, as is typically expected to be the case. To keep the relative number of points (i.e. $V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $\epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.1.

### Non-central slice {#sec:offcentre}

The equations above assume that the projection plane, and thus the slice, passes through 0, which will be the mean of centred data. This (centre point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too. However it is straightforward to generalize the equations to use any centre point, $\mathbf{c}$.

@@

The slice tour has been implemented in R [@rref] as a new display method `display_slice` in the `tourr` package [@tourr]. In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the centre point $\mathbf{c}$ is set by the `anchor` argument. By default `eps=0.1` and `anchor=NULL`, resulting in slicing through the mean. In addition the user can select the marker symbols for point inside and outside the slice. By default points in the slice are highlighted as `pch_slice = 20` and `pch_other = 46`, i.e. plotting a bullet for points inside the slice, and a dot for points outside. Below we show example code for displaying slices through a hollow 3D sphere, similar to the example in Figure \ref{fig:renderFunction}.

@@

It is especially instructive to look at spheres that are sliced off-centre. In this case the different views obtained in the slice tour reveal more of the concave structure of the distribution. The first row of Figure \ref{fig:3anchored} shows example views from a slice tour on a 3D sphere with $R=1$ and centre point $(0.7, 0.7, 0.7)$. Notice that this was chosen to fall outside the sphere. Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $R$ when the centre point has a negligible orthogonal component to the viewing plane (shown on the right); a circle with radius $<R$ as the viewing angle is tilted away from the axis connecting the centre point to the origin (shown in the middle); a full circle with small radius as the angle increases (shown on the left); and finally we see an empty slice when the projection plane is orthogonal to the centre point axis. A similar picture is found for the 5D sphere, see second row of Figure \ref{fig:3anchored}. For this example we generate 20k points on a hollow 5D sphere to resolve the features. Since $h$ increases with $p$ the resolution is reduced compared to the 3D example. Note also that as dimensionality increases, the larger orthogonal space means that the centre point will have a large orthogonal component in most views.

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored/sphere-3-anchored-033.png}
\hspace{5mm}
\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored/sphere-3-anchored-036.png}
\hspace{5mm}
\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored/sphere-3-anchored-041.png}}
\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-014.png}
\hspace{5mm}
\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-010.png}
\hspace{5mm}
\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-001.png}}
\caption{Different slices through a 3D (first row) and 5D (second row) hollow sphere with $R=1$, with shifted centre point, showing a full circle with small radius (left), a circle with radius $<R$ (middle) and a circle with radius $R$ (right).}
\label{fig:3anchored}
\end{figure*}

### Other geometric shapes
To better understand the slice tour we look at different examples of geometric shapes, see Figure \ref{fig:geoms}. For each shape two selected views are shown. The first column shows views from the slice tour on a 3D Roman surface. The second column shows slices through a 4D torus, revealing different aspects of the shape. The last column shows a 6D cube, where the upper plot shows a view along two of the original parameters, allowing to clearly identify the rectangular shape. The panel below shows that this is not typically the case when looking at a randomly selected slice.

@@

Selected views of projections and sliced projections from a radial basis SVM on 3 variables, and a polynomial basis SVM on 5 variables are shown in Figure \ref{fig:wine}. The slicing allows exploring the centre of the space. With 3D it reveals the spherical boundary of the group (green) hidden by the projection, as seen by comparing the top and bottom left plots of Figure \ref{fig:wine}. In 5D with the polynomial basis, projection might suggest that the boundary between classes is almost linear (top middle and right plots of Figure \ref{fig:wine}), but the slicing shows it to be nonlinear near the centre (bottom middle and right plots of Figure \ref{fig:wine}).

\begin{figure*}[ht]
\flushleft{\hspace{5mm} Projection, radial basis \hspace{25mm}  Projection, polynomial basis}
\centerline{
\includegraphics[width=0.3\textwidth]{gifs/png/wine-radial-projected/wine-radial-projected-031.png}
\hspace{5mm}
\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly-projected/wine-poly-projected-096.png}
\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly-projected/wine-poly-projected-042.png}}
\flushleft{\hspace{10mm} Slice, radial basis \hspace{30mm}  Slice, polynomial basis}
\centerline{
\includegraphics[width=0.3\textwidth]{gifs/png/wine-radial/wine-radial-031.png}
\hspace{5mm}
\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly/wine-poly-096.png}
\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly/wine-poly-042.png}
}
\caption{Exploring classification boundaries of the wine dataset using projection (top row) and sliced projection (bottom row). Radial basis SVM (first column) of 3 variables shows how the slicing reveals the spherical shape of the boundary of one group (green) that was hidden in the projection. Polynomial basis SVM (second and third columns) of 5 variables. The orange group does tend to be wrapped by the green group, and the blue group disappears with slicing, showing that it is on the outer edge of the space.}
\label{fig:wine}
\end{figure*}

@@

This paper has introduced a new visualization method for dynamic slicing of high-dimensional spaces. It is based on interpolated projections obtained in a (grand) tour and generates an interpolated sequence of sliced projections. The examples shown in Section \ref{sec:examples} demonstrate the potential of this new display to find and explore concave structures, as well as other hidden features. \textcolor{red}{Slice tour animations for all examples are available at \url{https://github.com/uschiLaa/paper-slice-tour}.}

@@

The approach by @prosection is more complex, and slices a subspace  of the $p-d=p-2$ dimensional space orthogonal to the projection. This generates more parameters, making it more difficult to navigate. More parameters mean more decisions on what to show. However, it is one of the next steps to explore different definitions of a slice tour, and the types of structure that might be captured by variations in slicing algorithms.

@@

The authors gratefully acknowledge the support of the Australian Research Council. The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr]. The source material and animated gifs for this paper are available at \url{https://github.com/uschiLaa/paper-slice-tour}. An appendix deriving the relative slice volume for the hypersphere is included as supplemental material. \textcolor{red}{We thank Andreas Buja for his contributions to the initial stage of this project.}
